A Guide to the Continuous Constant pH Molecular Dynamics Methods in Amber and CHARMM v[\versionnumber]

### 2022-01-23 cphmd_tutorial.tex
Initial Overleaf Import

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{hyperref} % Not original - Added by Jack
\usepackage{listings} % Not original - Added by Jack
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\newcommand{\pka}{p\textit{K}$_{\rm a}$}
\graphicspath{{figures/}}

@@

\newcommand{\versionnumber}{1.0}  
% you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/myaccount/homegithubrepository}}  %this should be the main github repository for this article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{A Guide to Continuous Constant pH Molecular Dynamics in AMBER and CHARMM v[\versionnumber]}

\author[1]{Jack A. Henderson}
\author[1]{Ruibin Liu}
\author[1,\authfn{2}]{Julie A. Harris}
\author[1,\authfn{3}]{Yandong Huang}
\author[1]{Vinicius M. de Oliveira}
\author[1*]{Jana Shen}
\affil[1]{University of Maryland School of Pharmacy, Baltimore, MD}
%\affil[2]{Institution 2}

@@

\orcid{Jack A. Henderson}{0000-0001-6675-7944}
\orcid{Ruibin Liu}{0000-0001-8395-9353}
\orcid{Julie A. Harris}{0000-0002-1130-6457}
\orcid{Yandong Huang}{0000-0002-1452-6383}
\orcid{Vinicius M. de Oliveira}{0000-0003-0927-3825}
\orcid{Jana Shen}{0000-0002-3234-0769}

% \contrib[\authfn{1}]{These authors contributed equally to this work}
% \contrib[\authfn{2}]{These authors also contributed equally to this work}

\presentadd[\authfn{2}]{Mapware, Washington DC}
\presentadd[\authfn{3}]{Jimei University, College of Computer Engineering, Xiamen, China}
% \presentadd[\authfn{4}]{Department, Institute, Country}

%\blurb{This LiveCoMS document is maintained online on GitHub at \githubrepository; to provide feedback, suggestions, or help improve it, please visit the GitHub repository and participate via the issue tracker.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Begin Abstract Here

\end{abstract}

@@

\section{Introduction} % - Vini

Variations in the pH condition of biological systems can lead to drastic changes in the physicochemical processes in these systems. 
Protein folding is one example of a natural pH-dependent process, where the unfolded protein only can reach its native state in a narrow pH range \cite{Pelegrine_Gasparetto_2005_LWT-FoodScienceandTechnology}. 
Thereby, small pH changes variation in the cell can lead to misfolding diseases and even cancer proliferation and metastasis \cite{vanderKamp_Daggett_2010_BiophysicalJournal,Webb_Barber_2011_NatRevCancer,Zheng_Liu_2020_TheInternationalJournalofBiochemistry&CellBiology}. 
The protonation state of residues is also critical in drug design once the ligand-protein binding is susceptible to charge changes \cite{Dominguez_Sussman_2010_Biochemistry,Bax_Edge_2017_ActaCrystallogrDStructBiol}. 
The SARS-CoV main protease (Mpro) is an example of a protein that possesses biological activity in a narrow pH range around 7.0 \cite{Tan_Hilgenfeld_2005_JournalofMolecularBiology}. Out of this range, its active site can collapse by protonation changes of residues that compose it \cite{Verma_Shen_2020_J.Am.Chem.Soc.}. More than that, the rational design of inhibitors must consider how the ligand interactions can alter the protonation state of the titratable residues into the binding site. The solution pH is also critical for the study, development, and fabrication of functional dynamic materials. An example of this is the biopolymer of chitosan, in which its physicochemical behavior is completely different at low and high pH conditions \cite{Morrow_Shen_2015_J.Am.Chem.Soc.}.

Classical molecular dynamics (MD) usually approach these pH-dependent systems by fixing the protonation states of titratable residues accordingly with experimental pKa values or pKa theoretical predictions. Several computational tools are currently used to predict pKa values; among the most used methods, we can cite the empirical ones PropKa and the Poisson-Boltzmann \cite{Olsson_Jensen_2011_J.Chem.TheoryComput.,Warwicker_Warwicker_2011_ProteinsStruct.Funct.Bioinforma.a}, and solver-based methods APBS-PDB 2PQR, DelPhiPKa, and H++ \cite{Unni_Baker_2011_J.Comput.Chem.,Wang_Alexov_2016_Bioinformatics,Anandakrishnan_Onufriev_2012_NucleicAcidsResearch}. Such tools can estimate the protonation states before running MD simulations, but they do not directly consider the dynamic flexibility of the molecules, which may lead to incorrect pKa predictions. Also, even assuming an accurate protonation states assignment at the beginning of MD, the conformational changes along the simulation can modify the electrostatic environment around titratable residues, affecting their protonation state. Thus, this relationship between conformational dynamics and the protonation states is not considered in the classical MD, and addressing this relationship is one of the most advantages of constant pH molecular dynamics (CpHMD) over conventional MD.

Over the past decades, much effort has been applied to developing and improving constant pH methods \cite{Chen_Shen_2014_Mol.Simulat.}. We can divide the most applied methods into discrete and continuous constant pH. The former periodically interrupts MD and uses Metropolis Monte Carlo (MC) criterion to sample the protonation states during simulation \cite{Burgi_Gunsteren_2002_ProteinsStruct.Funct.Bioinforma.}. In this case, the protonation of titratable groups is settled discretely, assuming one of the two states (protonated or deprotonated). On the other hand, the continuous CpHMD treats the protonation states of the ionizable sites with an auxiliary set of continuous titration coordinates, using an extended Hamiltonian $\lambda$-dynamics approach \cite{Chen_Shen_2014_Mol.Simulat.}. This method allows the system to escape local energy minima by accessing partially protonated states temporarily. For years, both MC and continuous CpHMD simulations have been performed and validated by pKa calculations for several systems in different environments \cite{Chen_Shen_2014_Mol.Simulat.,Radak_Roux_2017_J.Chem.TheoryComput.}. Also, CpHMD has been crucial to investigate the molecular mechanisms involved in the pH dependency of protein folding \cite{MartinsdeOliveira_PereiraLeite_2018_BiophysicalJournal,Yue_Shen_2018_Phys.Chem.Chem.Phys.}, enzyme catalysis \cite{Huang_Shen_2018_J.Phys.Chem.Lett.}, conformational changes \cite{Sarkar_Roitberg_2020_J.Phys.Chem.B}, protein/ligand binding \cite{Ellis_Shen_2016_J.Phys.Chem.Lett.,Harris_Shen_2017_J.Phys.Chem.Lett.}, activation of channels and transporters \cite{Huang_Shen_2020_}, and viral uncoating \cite{Chen_Shen_2016_J.Phys.Chem.Lett.}.

Our research group has exhaustively explored continuous CpHMD for different solvent models and force fields (FF). Using the GBNeck2 implicit solvent model implemented in Amber and hybrid-solvent based on the CHARMM platform, we accurately predicted pKa values for a large set of proteins \cite{Huang_Shen_2018_J.Chem.Inf.Model.,Harris_Shen_2019_J.Chem.Inf.Model.}. We also showed that these methods are reliable to predict cysteine and lysine protonation states, which are frequently used in covalent drug design \cite{Liu_Shen_2019_J.Am.Chem.Soc.,Harris_Shen_2020_J.Chem.TheoryComput.}. The active sites of SARS and MERS coronaviruses proteases were also investigated applying CpHMD, and we could clarify how proton-coupled conformational changes can alter their activity \cite{Verma_Shen_2020_J.Am.Chem.Soc.,Henderson_Shen_2020_J.Chem.Phys.}. Performing CpHMD simulations for Human Renin, we detailed its pH-dependent structure-dynamic-function relationship \cite{Ma_Shen_2020_J.Chem.Inf.Model.}, and explored the molecular mechanisms involved in inhibitors' binding and selectivity in BACE1 and $\beta$-Secretase \cite{Ellis_Shen_2015_J.Am.Chem.Soc.,Ellis_Shen_2016_J.Phys.Chem.Lett.,Harris_Shen_2017_J.Phys.Chem.Lett.}. Such examples are only a small sample of the vast application possibilities that CpHMD covers.

\begin{table*}[]
\begin{center}
\begin{tabular}{lp{0.5\linewidth}c}
System &  Topic  & CpHMD Method \\ 
\hline
% Add H. Line (Soluble)
BACE1   & Investigated pH-dependent ligand binding site dynamics, small molecule inhibitor interactions, and identify proton coupled allostery  & Hybrid-Solvent \\
BACE2   & Investigated how protonation state changes of a small molecule inhibitor affect binding with comparisons to BACE1 & Hybrid-Solvent \\
Cat-D   & Investigated pH-dependent ligand binding site dynamics and compared them to BACE1 & Hybrid-Solvent \\
Plasmepsin & Dissect the pH-dependent interaction of the ligand binding site, identify proton coupled allostery, elucidate the catalytic roles of the active site & Hybrid-Solvent \\
Renin   & Explore the acid/base roles of the catalytic dyad and conformational dynamics of the ligand binding site & Hybrid-Solvent \\
Silk Protein & & \\
SNase   &    & \\
Kinases &  Discovery Reactive Lys  & Implicit-Solvent \\
PLpro   &  Identify reactive Cys residues and elucidate proton-coupled conformational dynamics  & Implicit-Solvent \\
Mpro    & Explore protonation state dependent substrate binding site dynamics and identify reactive Cys resdiues & Implicit-Solvent \\
% Add H. Line (Subsection: Folding)
BBL     & & \\
NTL9    & & \\
% Add H. Line (Transmemb)
NhaA    & Addressed a controversy surrounding the proton binding residues involved in the sodium proton exchange mechanism & Membrane-Enabled Hybrid-Solvent \\
NapA    & Dissected the competitive ion binding mechanism and identified the proton binding sites. & Membrane-Enabled Hybrid-Solvent \\
AcrB    & & Membrane-Enabled Hybrid-Solvent \\
M2      & & Membrane-Enabled Hybrid-Solvent \\
\end{tabular}
\end{center}
\caption{Protein systems that CpHMD has been used on.}
\begin{minipage}{\columnwidth}
Possible Minipage
\end{minipage}
\label{Fixed-Charge_simulations}
\end{table*}

@@

%In this particular template, we have removed most of the usage examples which occur in \texttt{sample-document.tex} to provide a minimal template you can modify; however, we retain a couple of examples illustrating more unusual features of our templates/article class, such as the checklists, and information on algorithms and pseudocode.

@@

% This section gives the background and idea/purpose of CpHMD
% Importance of pH in simulations OK
% How Traditional MD Handles Titratable Sites OK
% How are titratable sites treated without pHMD (PropKA, Delphi,  H++, ect...) OK
% Idea of pHMD OK
% Brief description of types of pHMD (DpHMD and CpHMD) OK
% Advantages of CpHMD OK
% Examples of CpHMD in action should range from CHARMM Hybrid-Solvent CpHMD to Amber CpHMD simulations probing for reactive Cys and Lys targets.

% Figure Concept: A figure showing several different protein systems that we have investigated with CpHMD.

% Topics to Cover (Might Need it's own section):
%   1.) Brief Description of how titratable sites are handle, i.e. lambda dynamics and tautomers
%   2.) Give an overview of the extended Hamiltonian.
%   3.) Discuss the importance of pH-based replica exchange and its applications on CPUs and GPUs

% Figure Concept: Here would be a good place to show the classic MD protocol with the CpHMD portion and we could show a small example of the replica exchange process.

\subsection{Constant pH molecular dynamics}

In continuous CpHMD, we define a titration coordinate, $\lambda_i$, where $i$ represents the ionizable residues \cite{Huang_Shen_2018_J.Chem.Inf.Model.}. We attribute to this coordinate an implicit function of an underlying coordinate to assure $\lambda_i$ remains in a range between 0 and 1. Thus, the $\lambda_i$ expression is given by:

\begin{equation}
    \lambda_i = \sin^2(\theta_i).
\end{equation}

When $\lambda_i = 0$, the $i$th titratable group is fully protonated, and it is fully deprotonated when $\lambda_i = 1$. In simulation analysis, we usually apply cutoffs for protonated states ($\lambda^P < 0.1$), and for deprotonated states ($\lambda^U > 0.9$) \cite{}. After defined the titration coordinate, we can use an extended Hamiltonian to allow the simultaneous propagation of spatial (real) and titration (virtual) coordinates:

\begin{equation}
\begin{split}
\label{eq:Hamiltonian}
\mathcal{H}(\{r_a\},\{\theta_i\})=
&\sum_a\frac{1}{2}m_a\dot{r}_a^2+\sum_i\frac{1}{2}m_i\dot{\theta}_i^2+U^{\rm int}(\{r_a\})\\
&+U^{\rm hybr}(\{r_a\},\{\theta_i\})+U^{\ast}(\{\theta_i\}),
\end{split}
\end{equation}
where $a$ and $i$ represent the index for the spatial and titration-site coordinates, respectively. The two first terms of eq. \ref{eq:Hamiltonian} describe the kinetic energy for the atoms and the titration site. $U^{\rm int}$ is the titration-independent bond and non-bond energies from the non-titrating atoms. The fourth term, $U^{\rm hybr}$, depends on the solvent treatment. Herein, we will present the fully implicit solvent and the hybrid-solvent model, which $U^{\rm hybr}$ for both models includes Coulomb, vdW, and generalized Born (GB) energies and is written as

\begin{equation}
\label{eq:Uhybr}
U^{\rm hybr}(\{r_a\},\{\theta_i\})=U^{\rm elec}(\{r_a\},\{\theta_i\})+
U^{\rm vdW}(\{r_a\},\{\theta_i\})+U^{\rm GB}(\{r_a\},\{\theta_i\}).
\end{equation}

In eq. \ref{eq:Uhybr}, for the calculation of the Coulomb and GB electrostatic energies, we incorporate the atomic partial charges on the titrating residue and compute the linear interpolation between the values in the deprotonated and protonated states. Thus, we can write the charge on atom $j$, $q_j$, on the titrating residue $i$ as
%
\begin{equation}
q_j(\lambda) = (1-\lambda_i) q_j^{prot} +  \lambda_i q_j^{unprot}.
\label{eq:qlamb}
\end{equation}

The second term of eq. \ref{eq:Uhybr} represents the linear attenuation of the van der Waals (vdW) interactions between the titrating residue $i$ and its surrounding protein region:

\begin{equation}
U^{vdW} = (1-\lambda_i)U^{vdW}(i, p),
\end{equation}
where $p$ is the index of all other atom types, and $U^{vdW}$ is the protonation independent vdW interaction energy (Lennard-Jones 6-12 potential).

Finally, the energy component that only affects the titratable groups is given by the last term of eq. \ref{eq:Hamiltonian} and consists of three biasing potentials, 
\begin{equation}
\label{eq:biasingpotential}
U^{\ast}(\{\theta_i\})=-U^{model}(\lambda_i)+U^{barr}(\lambda_i)+U^{pH}(\lambda_i),
\end{equation}

The first term of $U^{\ast}$ describes a potential of mean force (PMF) along the $\lambda$-coordinate, considering $\lambda_i$ as an isolated model compound in the solvent,

\begin{equation}
\label{eq:Umod_1d}
U^{model}=A(\lambda_i-B)^2,
\end{equation}
where $A$ and $B$ are parameters that can be determined via thermodynamic integration (TI), which are performed by MD simulations fixing several $\lambda(\theta_i)$ values, and the parameters are obtained fitting the derivative of the mean force. The PMF is second-order in $\lambda$ and $x$ for residues with two tautomeric states, where $x$ describes the tautomeric state. In this case, the two-dimensional PMF can be written as a bivariate polynomial,
%
\begin{equation}
\begin{split}
U^{mod}(\lambda_i,x_i) =
&a_0 \lambda_i^2 x_i^2 + a_1 \lambda_i^2 x_i + a_2 \lambda_i x_i^2 + a_3 \lambda_i^2 \\
&+ a_4 x_i^2 + a_5 \lambda_i x_i + a_6 \lambda_i + a_7 x_i + a_8,
\end{split}\label{eq:double}
\end{equation}
where $x_{i}=\sin^{2}\left(\theta^{x}_{i}\right)$ where the $\theta^{x}_{i}$ are unbound titration variables that ensure that the $x_{i}$ are bound by 0 and 1.
%

$U^{barr}$, the second term of eq. \ref{eq:biasingpotential}, describes a barrier potential in the center of the titration coordinate, 
%
\begin{equation}
U^{barr}(\lambda_i) = -4\beta\left(\lambda_i - \frac{1}{2} \right)^2,
\label{eq:Ubarr}
\end{equation}
where $\beta$ is a parameter that determines the height of the barrier. Such potential increases the sampling time in the $\lambda$  coordinate endpoints and avoids unphysical vdW and electrostatic interactions associated with mixed titration state.

Finally, the last term of eq. \ref{eq:biasingpotential} is given by
%
\begin{equation}
U^{pH}(\lambda_i)=ln(10)k_BT(pH-\textrm{p}{K_a}^{mod})\lambda_i, 
\label{eq:UpH}
\end{equation}
% 
where $U^{pH}$ represents the pH-dependency of the deprotonation free energy, $k_B$ is the Boltzmann constant, $T$ is the system temperature, and $\textrm{p}{K_a}^{mod}$ is the model {\pka} for the titratable residue $i$.

\subsection{Scope} % - Ruibin 
This tutorial mainly covers system preparation, production run and restart, and post-processing data for the two mostly used CpHMD implementations, i.e., GBNeck2 implicit solvent based and GPU-accelerated implementation (GBNeck2-CpHMD) in the Amber pmemd.cuda platform, and the hybridge solvent and OpenMPI based implementation (Hybrid-CpHMD) in the CHARMM platform. For the GBNeck2-CpHMD version, we will first present how to prepare a system in explicit solvent like removing unwanted hetero atoms, ligands, ions, and waters, adding missing residues and atoms, adding dummy hydrogens, and post processing Amber parm7 file for updating exclusion list and changing GB radii. After that, we talk about how to pre-equilibrate the system before running CpHMD production simulations. For the production runs, we provide instructions on how to use conventional pH replica exchange on HPC and GPUs or asynchronous pH replica exchange on GPU clusters, and restarting existing jobs if necessary. For the Hybrid-CpHMD version, the system preparation process is smilar to GBNeck2-CpHMD but we mainly depend on CHARMM commands along with some mmtsb tools and we will learn how to add a pre-equilibrated water box, to miminize hydrogens with restraints and to head and equilibrate the whole system gradually. Instructions about how to use conventional pH replica exchange on HPC for Hybrid-CpHMD are followed as well. For all CpHMD simulations, choices about titratable residue types and pH conditions to use will be discussed for different use cases. At last, we will learn how to post process CpHMD data or the lambda files, and how to interpret the data with examples and explanations. We won't try to cover much traditional molecular dynamics analysis techniques but necessary solvent accessible surface area (SASA) and hydrogen bond analysis tools will be used to explain {\pka} upshifting or downshifting results. For all the topics, we will provide open-sourced bash or python scripts we used daily in our research.
% This section needs to describe what is covered in this review.
% Introduce CpHMD in the AMBER and CHARMM Packages, State the Types of CpHMD in AMBER and CHARMM, i.e. AMBER: Async pHREX with GBNeck2 and PME on GPUs, CHARMM: pHREX Hybrid-Solvent CpHMD on CPUs
% 1.) How to prepare a system for CpHMD
% 2.) Running CpHMD in CHARMM Equilibration to Production
% 3.) Running CpHMD in AMBER Equilibration to Production
% 4.) Analyzing Lambda Files, all the classic lambda values analysis procedures.

%Tutorials should endeavor to cover the specific task at hand, and also highlight how the steps might need to be modified (or additional care might need to be taken at particular points) to handle more general cases.

%The scope of the tutorial, as well as the expected proficiencies / outcomes for researchers who complete the tutorial, should be clearly defined.
%This will often happen in a specific section or subsection in the article itself.

We also have a GBSW implicit solvent based and a PME/GRF all-atom based implemenations in the CHARMM platform, but users are not encouraged to use them for speed issues. We are currently developing a GPU-accelerated all-atom implementation in the Amber pmemd.cuda platform, and the readers are welcomed to contact us for usage when it is finally finished.

\section{Prerequisites} % - Ruibin
CpHMD is based on classical molecular dynamics (MD) simulations, so it is a requirement that the user has some knowledge of MD and experience in running some MD jobs, especially in the Amber or CHARMM platform. Users with experience in other platforms like Gromacs, NAMD, or OpenMM are encouraged to go through some introductory tutorials for Amber or CHARMM to get started. For running a CpHMD simulation, it is important to know your system and research goal. If the user wants to know the {\pka} values for protein residues or their protonation states at certain pH conditions, the GBNeck2-CpHMD version is the best choice and Amber experience is required. If the user wants to investigate detailed proton-coupled conformational transormation, ion transfer, or ligand-protein binding/unbinding, especially when membrane system is involved, the Hybrid-CpHMD is currently the only choice and CHARMM experience is necessary. Like all computational chemistry calculations, it is important to know your system as much as possible before any serious simulations. With them, we can decide what salt concentration, temperature, pH range, and titratable residue types to use in CpHMD. Also, an expectation of time scale of the interested biological process is important to decide how long is required to run the CpHMD job.
% Who is the tutorial designed for, beginners or people with MD experience?
% What knoweledge should the user have prior to running CpHMD?
% Why should a person choose to run a simulation with CpHMD
% What should a person know before running a CpHMD simulation, pH-range, important titratable site... ect..

%Here you would identify prerequisites/background knowledge that are assumed by your work, as well as any software/license requirements.

\subsection{Background knowledge}
Knowledge of classical MD is required to understand what we can and cannot achieve through a MD simulation. Concepts in MD like force field, equation of motion, time step, convergence, thermostat, barostat, and so on are important to understand why we need them and the parameters we choose in our minimization, heating, equilibrating, and production runs. Of course, chemistry knowledge of pH and its significance in chemical, material, or biological systems are essential in entering this field.

Both the GBNeck2-CpHMD in Amber and Hybrid-CpHMD in CHARMM are primarily for Linux especially for HPC environments. The users are expected to have a good knowledge to use Linux command line tools for accessing and editing files, submitting jobs to servers, and retrieving simulation result files from remote Linux machines. Bash and/or Python scripting skills are required for pre- and post- processing files. For HPC environments, it is required to know the specific job submission and management system like SLURM or SGE installed in the HPCs. For GPU platform, knowledge of advantages of GPU acceleration in molecular dynamics and using tools such as 'nvidia-smi' to check GPU status and 'nvcc --version' to check CUDA version is necessary as well. The user can use any data visualisation tool(s), but matplotlib in Python and Jupyter notebooks is the first choice in our tool sets.  
% Tutorials should clearly define what concepts or abilities researchers will need to complete the tutorial (e.g., some proficiency in Python; experience with Jupyter notebooks; knowledge of classical MD; etc).

\subsection{Software/system requirements} % - All
% CHARMM Package Version, number of processors recommended for peak performance 
% AMBER Package version, Recommended GPUs
For GBNeck2-CpHMD in Amber, Amber 16 is the minimum version. Requirements for installing Amber can be found in the \href{https://ambermd.org/Installation.php}{Amber website}. For the GPU version, or specifically pmemd.cuda, the Nvidia GPU cards with at least compute compatibility sm3.0 should be installed with the latest driver. CUDA 10.1 is the primary CUDA version for running it but early versions from CUDA 7.5 also works. After successfully installing CUDA, it is important to set three environmental variables in Linux. CUDA\_HOME and LD\_LIBRARY\_PATH must point to the directories where the desired CUDA version is installed. We also need to specify CUDA\_VISIBLE\_DEVICES to GPU device indices. For example, 
\begin{lstlisting}
export CUDA_HOME=/usr/local/cuda-10.1
export LD_LIBRARY_PATH=\
            /usr/local/cuda-10.1/lib64
export CUDA_VISIBLE_DEVICES=0,1
\end{lstlisting}
will direct the Linux system to use the CUDA 10.1 version installed in the /usr/local directory and GPU with indices 0 and 1 will be used in calculations. After installing Amber, we should set the AMBERHOME environmental variable as the Amber package installation folder. 
\begin{lstlisting}
export AMBERHOME=/path/to/Amber
\end{lstlisting}

To prepare systems for CpHMD simulations, we usually use \href{http://blue11.bch.msu.edu/mmtsb/Main_Page}{mmtsb tool set} for CHARMM and our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-prep}{cphmd-prep} tools for Amber. We also use a script-level \href{https://gitlab.com/shenlab-amber-cphmd/async_ph_replica_exchange}{asynchronous replica exchange} implementation in our lab or GPU pH replica exchange simulations. Scripts in the \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-analysis}{cphmd-analysis} are used for analyzing CpHMD lambda files, including convergence checking, {\pka} calculation, and replica profiles. Some simple trajectory based analysis like RMSD can also be done but the users can use any tools from cpptraj, pytraj, vmd, mdanalysis, and more. Note that these free tools usually depends on other free softwares, including OpenMM, Ambertools, scipy, pandas, matplotlib, etc. The instructions on how to install those softwares are covered in their individual pages and not discussed here. A general suggestion to use Python depending packages is to use Anaconda or Miniconda. With them, the user can create an environment specific for CpHMD related calculations. For example, after installing Anaconda or Miniconda, we can type 
\begin{lstlisting}
conda create --name cphmd python=3.7
\end{lstlisting}
Note that we choose Python version 3.7 because the OpenMM package currently only supports up to Python 3.7. Activate the environment named 'cphmd' by
\begin{lstlisting}
conda activate cphmd
\end{lstlisting}
Many packages can be installed through the conda and pip tool.
\begin{lstlisting}
conda install -c omnia pdbfixer parmed
conda install -c conda-forge ambertools \
                 mdanalysis matplotlib 
conda install -c anaconda numpy scipy pandas 
pip install f90nml
\end{lstlisting}
Whenever the user wants to do some calculations or analysis related to CpHMD, they can activate this 'cphmd' environment and use the installed tools without conflicting with other software. 
% Python Packages and Version required for Async-Replica Exchange 
% Python Packages and Version required for the CpHMD-Analysis Lambda Parser

%Tutorials should clearly define what system and/or software requirements the researcher will need to complete the tutorial (e.g., VMD version 1.9 or newer, AMBER, etc.). Tutorials requiring specific software packages must provide instructions and files for the referenced version of the software.

% \section{Content and links}

% A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
% You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.

% \section{Checklists}
% Tutorials do not necessarily require the use of a checklist as in Best Practices documents; however, they can include these if desired.
% Several useful checklist formats are available, with examples presented in \texttt{sample-document.tex} in \url{github.com/livecomsjournal/article_templates/templates}.
% One example is shown here.

\section{Workflow for CpHMD simulations} % - Jack/Ruibin
% State an outline of the steps necessary for running the CpHMD simulations in general.

% Figure concept: Show a diagram of the steps needed to run a CpHMD simulation and images representing those steps.

 %--------------------------------------------- 
\begin{figure*}
    \centering
    \includegraphics[width=6.6in]{figs/workflow_scheme.pdf}
    \caption{\textbf{Generalized CpHMD workflow} 
    *Mock Figure*  
    }
    \label{fig:intro}
\end{figure*}
 %---------------------------------------------

\subsection{Hybrid-Solvent CpHMD in CHARMM} % - Jack

Here, the steps and processes involved in running CpHMD simulations in CHARMM are presented. 
In the first part of each section (System Preparation, Equilibration, and Production) an example of how to run CpHMD on the small soluble protein BBL (PDBid: 1w4h)\cite{Ferguson_JMolBiol_2005_v353_p427} is used as an example case. 
To follow along the please download the CpHMD tutorial from our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/pH-REX_CpHMD_BBL_tutorial_in_CHARMM}{GitLab page}.
In the second part of the system preparation and equilibration section, supplemental information is provided discussing how to setup and run CpHMD simulation on transmembrane proteins.
Example input files for select portions of this process can be found from our \href{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial/-/tree/main/MEMB_pH-REX_CpHMD_CHARMM_Examples}{GitLab page}.

\subsubsection{System Preparation}

% Figure Concept: Show image of the ASP in the double protonated (syn-Configuration) State

>>>>

\usepackage{lipsum} % Required to insert dummy text
\usepackage[version=4]{mhchem}
\usepackage{siunitx,bm}
\usepackage{hyperref} % Not original - Added by Jack
\usepackage{listings} % Not original - Added by Jack
\DeclareSIUnit\Molar{M}
\usepackage[italic]{mathastext}
\newcommand{\pka}{p\textit{K}$_{\rm a}$}
\graphicspath{{figures/}}

@@

\newcommand{\versionnumber}{1.0}  
% you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://gitlab.com/shenlab-amber-cphmd/cphmd-tutorial}}  %this should be the main github repository for this article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{A Guide to the Continuous Constant pH Molecular Dynamics 
Methods in Amber and CHARMM v[\versionnumber]}

\author[1,\authfn{1},\authfn{2}]{Jack A. Henderson}
\author[1,\authfn{1}]{Ruibin Liu}
% \alsoaffiliation{
% Joint first author
% }
\author[1,\authfn{3}]{Julie A. Harris}
\author[1,\authfn{4}]{Yandong Huang}
\author[1]{Vinicius M. de Oliveira}
%\author[1]{Paween Mahinthichaichan}
\author[1*]{Jana Shen}
\affil[1]{University of Maryland School of Pharmacy, Baltimore, MD}
%\affil[2]{Institution 2}

@@

\orcid{Jack A. Henderson}{https://orcid.org/0000-0001-6675-7944}
\orcid{Ruibin Liu}{https://orcid.org/0000-0001-8395-9353}
\orcid{Julie A. Harris}{https://orcid.org/0000-0002-1130-6457}
\orcid{Yandong Huang}{https://orcid.org/0000-0002-1452-6383}
\orcid{Vinicius M. de Oliveira}{https://orcid.org/0000-0003-0927-3825}
\orcid{Jana Shen}{https://orcid.org/0000-0002-3234-0769}

\contrib[\authfn{1}]{These authors contributed equally to this work}
% \contrib[\authfn{2}]{These authors also contributed equally to this work}

\presentadd[\authfn{2}]{Scorpion Therapeutics, Boston, MA}
\presentadd[\authfn{3}]{ComputChem LLC, Baltimore, MD}
\presentadd[\authfn{4}]{Jimei University, College of Computer Engineering, Xiamen, China}
% \presentadd[\authfn{4}]{Department, Institute, Country}

\blurb{This LiveCoMS document is maintained online on GitLab at \githubrepository. To provide feedback, suggestions, or help improve it, please visit the GitLab repository and participate via the issue tracker.}

@@

% \lstnewenvironment{listings}[1][] 
%  {\lstset{frame=off,escapechar=`,breaklines, breakindent=10pt, breakatwhitespace=True, linewidth=3.4in, #1}}
% {}

\lstnewenvironment{plaincontent}[1][] 
 {\lstset{frame=shadowbox,escapechar=`,breaklines, breakindent=10pt, breakatwhitespace=True, linewidth=3.4in, #1}}
{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Begin Abstract Here
Like temperature and pressure, solution pH is an important environmental variable in biomolecular simulations. 
Virtually all proteins depend on pH to maintain their structure and function. 
In conventional molecular dynamics (MD) simulations of proteins, pH is implicitly accounted for by assigning and fixing protonation states of titratable sidechains. This is a significant drawback, 
as the assigned protonation states may be wrong and they may change during dynamics.
In this tutorial, we guide the reader in learning and using the various continuous constant pH MD methods in Amber and CHARMM packages, which have been applied to predict {\pka} values and elucidate proton-coupled conformational dynamics of a variety of proteins including enzymes and membrane transporters.
\end{abstract}

@@

@@

@@

\input{1_intro}
\input{1_methods}
\input{1_scope}
\input{2_prereq}
\input{3_hphmd}
\input{4_aphmd}
\input{5_gphmd}
\input{6_analysis}
\input{7_parameterization}

\section{Author Contributions}
%%%%%%%%%%%%%%%%
% This section mustt describe the actual contributions of
% author. Since this is an electronic-only journal, there is
% no length limit when you describe the authors' contributions,
% so we recommend describing what they actually did rather than
% simply categorizing them in a small number of
% predefined roles as might be done in other journals.
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io
% for more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%%
Henderson, Liu, Harris, Huang, de Oliveira, and Shen
together wrote the manuscript.
We thank Zhi Yue and Paween Mahinthichaichan for contributing example scripts for the hybrid-solvent CpHMD simulations of transmembrane proteins.
We thank Jason A. Wallace for a bash script to calculate {\pka} values and track the convergence of unprotonated fractions using Grace.
% % We suggest you preserve this comment:
% For a more detailed description of author contributions,
% see the GitHub issue tracking and changelog at \githubrepository.

%\section{Other Contributions}
%%%%%%%%%%%%%%%
% You should include all people who have filed issues that were
% accepted into the paper, or that upon discussion altered what was in the paper.
% Multiple significant contributions might mean that the contributor
% should be moved to authorship at the discretion of the a
%
% See the policies ``Policies on Authorship'' section of https://livecoms.github.io for
% more information on deciding on authorship and author order.
%%%%%%%%%%%%%%%

% (Explain the contributions of any non-author contributors here)
% % We suggest you preserve this comment:
% For a more detailed description of contributions from the community and others, see the GitHub issue tracking and changelog at \githubrepository.

\section{Potentially Conflicting Interests}
%%%%%%%
%Declare any potentially competing interests, financial or otherwise
%%%%%%%
The CpHMD-related software may be of interest to ComputChem LLC, for which
J.S. is a founder and scientific advisor. 
J.S. also serves on the scientific advisory board of EoCys LLC.

\section{Funding Information}
%%%%%%%
% Authors should acknowledge funding sources here. Reference specific grants.
%%%%%%%
We acknowledge financial support from the National Institutes of Health (R01GM098818 and R01CA256557 to Shen and R44GM134756 to Harris) and National Science Foundation (CBET1932963 to Shen).
